The importance of measurement uncertainty in ecological management

### 2018-08-17 manuscripts/manuscript.tex
nearly there

Imperfect information is ubiquitous in ecological management and
conservation decision making. While the pressing concerns of global
change have put the spotlight on \emph{forecasting}, i.e.~the
uncertainty of the future (e.g. Petchey et al. 2015), management
decisions must also contend with uncertainty of the \emph{present}: How
many fish are in the sea today? What regions harbor the most fragile
biodiversity? How have invasive species transformed native communities?
A recognition of the importance of uncertainty has become widespread in
the context of estimating ecological models from available data, thanks
in large part to the rise of approaches such as Hierarchical Bayesian
Modeling, along with convenient software tools for implementing these
methods. In sharp contrast to that shift, the treatment of uncertainty
in decisions -- how the uncertainty in models should be reflected in
scientific recommendations, remains far less developed, more
qualitative, and under-served by available tools. We illustrate how
uncertainty in measurement has a dramatic negative impact on outcomes
under a range of existing approaches, which can be avoided through the
use of more formal approaches that have since been refined in fields
such as robotics. This insight both resolves a long-standing paradox in
fisheries management and has important implications for how ecosystem
management deals with uncertainty more generally.

Approaches for decision making under uncertainty in ecological systems
can be divided into two camps: (1) those based upon ``optimal control''
solutions, and usually favored by natural resource economists, and (2)
those based in heuristic methods such as scenario planning, resilience
thinking, and precautionary rules of thumb, more commonly found in both
ecological literature and actual practice (Fischer et al. 2009; Polasky
et al. 2011). While researchers have recognized the need to unify the
transparent and quantitative algorithmic approach of optimal control
with the greater complexity and uncertainty of real ecosystems that is
acknowledged by heuristic methods, computational barriers to doing so
have stymied this progress. We illustrate how the limitations of these
approaches have manifested in the example of fisheries management, and
present a new approach that can combine the reality of measurement
uncertainty and the rigor of optimization to resolve a long-standing
paradox and suggest a more robust approach to management. We compare the
results of managing simulated fish stocks under (A) the existing optimal
control approach, (B) a heuristic approach that is more typical of
today's management practices, and (C) a newly proposed approach that
provides a more rigorous treatment of uncertainty.

The challenge of making decisions under uncertainty is obviously not
unique to fisheries, and the existing methods we critique here can be
found in a wide range of population and ecosystem management challenges,
including fire ecology (Richards et al. 1999), invasive species
({\textbf{???}}; Blackwood et al. 2010), disease outbreaks (Shea et al.
2014; Li et al. 2017), and protecting biodiversity (Dee et al. 2017;
Iacona et al. 2017). Yet fisheries conservation and management has long
been both a crucible and proving ground for the theory of ecological
management more generally, including topics such as adaptive management
(Walters and Hilborn 1978), ecosystem-based management (Levin and
Lubchenco 2008) and resilience thinking (Holling 1973; May 1977), while
also giving rise to the sub-discipline of resource economics (Gordon and
Press 1954; Schaefer 1954; Beverton and Holt 1957), thanks to its global
relevance, long history, and readily available data. Despite the
generality of the models and issue of measurement uncertainty considered
here, any application must be grounded in both the history and context
of the particular decision problem. In light of this, we begin with a
brief background on existing decision theory approaches common in
fisheries management.

@@

For simplicity we will assume measurement uncertainty is normally
distributed with standard deviation \(\sigma_m\),
\(\xi \sim \mathcal{N}(1, \sigma_m)\). Given an estimated recruitment
model along with this measurement of the stock size, the manager must
choose the set of a harvest quotas \(\lbrace h_t \rbrace\) for
\(t \in [0, 1, \infty]\) that maximizes expected long term (net) present
value:

\begin{equation}
\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}
\end{equation}

A few subtleties arise in this seemingly simple problem statement that
we must address. First, given our discrete-time formulation of the
recruitment process, it is necessary to decide if the measurement
\(y_t\) happens before or after harvest \(h_t\): that is do we: measure,
recruit, harvest, or measure, harvest, recruit? Following convention in
the optimal control literature (e.g. (Reed 1979; Clark 1990; Sethi et
al. 2005)), we will assume the latter; measurement occurs immediately
before harvest, followed by recruitment.

A second subtly is in the nature of the decision problem itself: note
that the manager in not myopic, thinking only about the next year, but
rather considers the sequence of all future actions \({h_t}\). While the
manager will revisit this sequence each year in light of new
observations, it is this ability of the decision problem to think ahead
that allows it to tolerate short-term costs (e.g.~reduced harvests in
the current season) for larger future payoffs. In this, ecological
management is a game of chess, always thinking several moves ahead. This
differs from the problem solved in theory as well as the practice of
maximum sustainable yield, which seeks to identify only a fixed
mortality \(F_{MSY}\) applied for all time, rather than a \emph{policy}
of varying quotas \(h_t\) depending on the stock assessment \(y_t\).
Understanding this distinction is key both to understanding the
differences between current theory and current best practice in
fisheries management.

@@

It is important to remember that both existing methods we discuss and
the POMDP approach proposed here can or are already applied to more
complex models as well, including those with age or stage structure.
Likewise, while we consider the model and its parameters as ``given,''
it should be understood to mean that this in practice comes as the
result of some model choice and statistical estimation from historical
data, frequently accounting for uncertainty in both the process and the
measurements (e.g. Dichmont et al. 2016). As we shall see here, having
included uncertainty in the model \emph{estimation} in no way excuses us
from having to also deal once again with that uncertainty in the
decision process.

@@

It is important to note that actual management strategies are quite
diverse. Our focus on comparisons TAC, MSY, and Constant Escapement (CE,
also refereed to as MEY, see below) reflects only the subset of current
practices. Another increasingly common approach to address uncertainty
in both process and models is known as Management Strategy Evaluation
(MSE, Smith 1994; {\textbf{???}}; Bunnefeld et al. 2011; Punt et al.
2016). MSE performs forward simulations under a suite of pre-selected
candidate strategies to determine which strategy achieves the best
expected outcomes. Crucially, by limiting the analysis to a suite of
candidate strategies over a finite number of replicate simulations, this
approach is able to accommodate substantially greater complexity and
uncertainty than has been possible with more formal optimal control
approaches we discuss below. Optimal control solutions rely on dynamic
programming to consider all possible strategies (all possible sequences
of harvest under all possible realizations of stock dynamics). As such,
MSE can be viewed as an approximation to the optimal control solutions
considered below, a notion that has been formalized in the literature on
approximate dynamic programming (Nicol and Chadès 2011).

@@

In the face of a naturally fluctuating population, previous arguments
made by Gordon and Press (1954); Schaefer (1954) or Clark (1973) no
longer hold. The optimal harvest of a population under a model such as
the stochastic Gordon Schaefer, Eq \eqref{Gordon-Schaefer}, is an
example of a Markov Decision Problem (MDP). This is not to be confused
with the the problem of \emph{estimating} parameters or posteriors of a
Markov process model, e.g.~through typical Bayesian analysis -- recall
this is a decision problem, in which the model has already been
estimated and now we are tasked with choosing a sequence of future
harvests based on that model. Such MDP problems are commonly referred to
in ecological literature by its solution method, Stochastic Dynamic
Programming (SDP) (Marescot et al. 2013), widely used in both ecological
management and animal behavior (Mangel 1985; Mangel and Clark 1988).
However, thanks to a mathematical proof provided by Reed (1979),
fisheries managers have largely been able to bi-pass the need for SDP.
But this convenient result also opened the door to paradox of
uncertainty that would persist for the next four decades.

@@

In practice, TAC-managed fisheries are more common that CE-managed
fisheries, at least in the United States. CE-based regulation is limited
primarily to salmon, species for which population estimates may be more
precise and thus closer to Reed's assumption. We are not aware of any
fishery that has put Clark's Paradox into practice, harvesting more in
response to increased uncertainty. As we shall illustrate, this is
particularly fortunate; though approaches that properly account for both
forms of uncertainty can perform even better.

@@

Why does the introduction of measurement uncertainty make the decision
problem so much harder? Consider the thought experiment of managing a
fish stock where in each of the previous four years stock assessments
have put the population size at, say, 51, 54, 49, 51, (say, measured in
percent of carrying capacity, \(K\)). Under a constant escapement target
of 50, the harvest quota would be 1, 4, 0, 1, respectively. Note that
this control rule depends only on the measurement for the year in
question, which is assumed to be measured without error. If we then
measure a stock size of 75, an assumption that this measurement is made
without error would justify the sudden increase to a quota of 25. Yet if
we admit the possibility of measurement error, this sudden bump starts
to look very suspicious in light of all the previous observations.
Perfect measurements let us exploit the Markov property of the model --
the most recent measurement tells us all we need to know. In contrast,
the introduction of measurement uncertainty breaks this Markovian
assumption: a rational decision maker would suspect this 75 to be an
over-estimate based on prior observations. Under imperfect observations,
we no longer have a Markov Decision Process (MDP) over the observed
state space. Rather, we have a Partially Observed Markov Decision
Process (POMDP). A POMDP policy cannot be expressed as a policy function
at all (much less as policy function with a single constant-escapement
target, as much previous work has assumed). Instead, a POMDP policy can
be thought of as depending not only on the most recent observation, but
the recent observation combined with a \emph{prior belief} about what
state the system is in, which itself is determined by all past
observations. In this manner, the POMDP can consider how much weight to
give the most recent observation based on how it compares to prior
beliefs.

@@

@@

We consider the average fish biomass across 100 replicate simulations
under three different management strategies: constant escapement (CE),
total allowable catch (TAC; equal to 80\% MSY), and partially observed
Markov decision process (POMDP) management. Each suite of simulations is
considered under three sequentially higher measurement error regimes,
from \(\sigma_m = 0\), \(\sigma_m = 0.1\), and \(\sigma_m = 0.15\), as
indicated, while stochastic recruitment (environmental noise) is set to
a moderate \(\sigma_g = 0.15\). In the absence of either environmental
noise or measurement error POMDP, CE, and MSY would converge to stock at
the \(B_{MSY}\), while TAC would maintain the stock at a slightly higher
level (see Supplemental material, Figure S1). The introduction of
stochastic growth has a significant negative impact on the TAC strategy,
with a mean biomass significantly lower than \(B_{MSY}\) (first panel,
Figure \ref{sims}; this impact is even more severe for MSY, see Figure
S2). Without measurement error, the CE and POMDP strategies are nearly
identical with both approximately maintaining the stock at \(B_{MSY}\)
despite the significant environmental stochasticity. As measurement
error increases in the subsequent panels, CE and TAC strategies perform
increasingly poorly, while the POMDP continues to maintain the average
stock close to \(B_{MSY}\). Notably, CE is significantly more impacted
by measurement uncertainty than TAC, with CE averaging even lower
biomass than TAC under moderate measurement uncertainty. These results
confirm that the precautionary approach represented by the TAC does
indeed prove more robust to the problem of measurement error than the
optimization solution represented by CE, but is less robust to
environmental error. In contrast, the POMDP approach successfully
handles both sources of uncertainty, maintaining higher stocks, near
\(B_{MSY}\) level, and producing the highest yield.

@@

Using modern algorithms we have been able to both crack the nut of
measurement uncertainty in the optimal management of ecological
populations such as marine fisheries, debunking a long-standing paradox
created by previous literature which has argued that increased
measurement uncertainty should be met with equal or larger harvests
(e.g. Reed 1979; Clark and Kirkwood 1986; Sethi et al. 2005). By using
powerful algorithms developed first the robotics literature to solve
POMDPs (Kurniawati et al. 2008), we have been able to tackle this
optimal control problem directly rather than resorting to the
simplifying assumptions used in previous work. Our solution confirms the
initution of the precautionary principle in a quantitatively precise
form: under most (but not all) conditions, \emph{greater uncertainty
results in smaller optimal harvests}. Despite frequently proposing
smaller harvests than existing approaches, the POMDP solution achieves
higher long-term economic value as well as higher average stock sizes
over the long term.

This lesson has clear relevance in the context of fisheries management.
Here, practice is already wiser than theory, where the optimal control
based on the CE results of Clark (1973) and Reed (1979) is limited to
fisheries such as salmon where estimates in stock size may be less
significant than in purely marine species. To our knowledge, no fishery
has intentionally put the more aggressive harvest solutions of Clark and
Kirkwood (1986) and Sethi et al. (2005) into practice. Yet this does not
lift the spectre of these methods, where CE-based predictions feature
prominently in global fisheries analyses as the economically optimal
strategy and the fastest route to recovery (Costello et al. 2016;
Burgess et al. 2018). Our results raise serious questions about the
optimistic forecasts predicted under these CE-based policies. Meanwhile,
MSY remains the dominate standard international law (Mace 2001), and
many well-managed US fisheries rely on a version of the TAC based
precautionary adjustment to MSY to comply with catch limits under
Magnuson-Stevens Act. Our results show that these strategies are less
sensitive to measurement error, but not immune. Comparisons against more
heuristic decision methods are more difficult: management strategy
evaluation (MSE; Smith 1994; Punt et al. 2016) will depend on precisely
what management scenarios are considered. Under sufficient uncertainty,
both the existing optimal and precautionary approaches can lead to
declines in biomass and economic returns which could be avoided by
POMDP-based management.

This conclusion has important broader implications for ecological
management more generally. Decision making under uncertainty is a
central challenge for ecology and conservation biology, which frequently
pits optimal control based approaches against more heuristic approaches
such as resilience thinking, scenario planning, and the precautionary
principle (Polasky et al. 2011). With economists frequently favoring the
precise, stylized approach of optimal control and ecologists embracing
the realities of greater complexity and uncertainty echoed in more
heursitic approaches, this division can sometimes appear to pit economic
outcomes against ecological objectives. That perspective is echoed in
the paradox discussed here, in which economic optimization ignoring
uncertainty has argued in cold mathematical logic for a rejection of the
precationary principle. Our results demonstrate that this divide is
artificial. With sufficiently powerful algorithms we can go some ways to
realizing the call of Fischer et al. (2009) to combine more ecologically
realistic assumptions such as measurement error with optimal control
approaches in ecological management. Both sides are vindicated in our
solution: economic optimization under POMDP does not controdict the
precautionary principle, while under POMDP, optimal control can also
achieve better ecological outcomes than blunt precautionary rule
(i.e.~TAC, which reduced all harvests by 20\% below MSY target). As
Fischer et al. (2009) predicted, unifying optimal control with more
resilience-inspired approaches is win-win.

If accounting for measurement uncertainty can resolve previous paradoxes
and divides, it also raises additional challenges. The approach
discussed here is vulnerable both to the charge of being too simple to
be realistic while also being too complicated to be feasible. These
issues are as easily recognized in other ecological management contexts,
from harvests of wild game or forestry to pest outbreaks, (e.g. Richards
et al. 1999; {\textbf{???}}; Blackwood et al. 2010; Shea et al. 2014;
Dee et al. 2017; Iacona et al. 2017; Li et al. 2017) as in fisheries,
where ecological complexity is invariably high and management capacity
invariably limited. We address each of these in turn.

>>>>

Imperfect information is ubiquitous in ecological management and
conservation decision making. While the pressing concerns of global
change have put the spotlight on forecasting, i.e.~the uncertainty of
the future (e.g. Petchey et al. 2015), management decisions must also
contend with uncertainty of the present: How many fish are in the sea
today? What regions harbor the most fragile biodiversity? How do we make
use of a rapidly expanding but opportunistically collected data, such as
in citizen science efforts, which may involve strong sampling biases?
Much of our focus in dealing with uncertainty has been through the lens
of statistical approaches to model fitting. The rise of approaches such
as Hierarchical Bayesian Modeling have given us sophisticated tools to
reflect uncertainty in our data and uncertainty in our knowledge of
mechanisms and parameters in our models (Ellison 2004). But estimating a
model is not the same as making a decision, particularly when a model
leaves us with considerable uncertainty about current and future states.
A model can only provide the rules of the chess board. It is the role of
decision theory to provide us with a strategy.

Decision theory has a long history in the ecological literature (e.g
Schaefer 1954; Walters and Hilborn 1978; Clark 1990; Shea 1998; Polasky
et al. 2011) which has sometimes been overshadowed by recent emphasis on
model fitting. While in principle a model will be updated throughout the
decision process (as we will discuss), it is convenient to think of the
model as ``given,'' estimated, with uncertainty, using the best
available methods. The goal of a decision theory is not maximize some
abstract notion of model fit (least squares, likelihood, etc), but a
more practical notion of utility (happiness, economic value, or any
combination of stakeholder objectives, see Halpern et al. (2013)).
Approaches for decision making under uncertainty in ecological systems
can be divided into two camps: (1) those based upon ``optimal control''
solutions, and usually favored by natural resource economists, and (2)
those based in heuristic methods such as scenario planning, resilience
thinking, and precautionary rules of thumb, more commonly found in both
ecological literature and actual practice (Fischer et al. 2009; Polasky
et al. 2011). While researchers have recognized the need to unify the
transparent and quantitative algorithmic approach of optimal control
with the greater complexity and uncertainty of real ecosystems that is
acknowledged by heuristic methods, computational barriers to doing so
have stymied this progress. We illustrate how the limitations of these
approaches have manifested in the example of fisheries management, and
present a new approach that can combine the reality of measurement
uncertainty and the rigor of optimization to resolve a long-standing
paradox and suggest a more robust approach to management.

The challenge of making decisions under uncertainty is obviously not
unique to fisheries, and the existing methods we critique here can be
found in a wide range of population and ecosystem management challenges,
including fire ecology (Richards et al. 1999), invasive species (Shea
1998; Blackwood et al. 2010), disease outbreaks (Shea et al. 2014; Li et
al. 2017), and protecting biodiversity (Dee et al. 2017; Iacona et al.
2017). Yet fisheries conservation and management has long been both a
crucible and proving ground for the theory of ecological management more
generally, including topics such as adaptive management (Walters and
Hilborn 1978), ecosystem-based management (Levin and Lubchenco 2008) and
resilience thinking (Holling 1973; May 1977), while also giving rise to
the sub-discipline of resource economics (Gordon and Press 1954;
Schaefer 1954; Beverton and Holt 1957), thanks to its global relevance,
long history, and readily available data. Despite the generality of the
models and issue of measurement uncertainty considered here, any
application must be grounded in both the history and context of the
particular decision problem. In light of this, we begin with a brief
background on existing decision theory approaches common in fisheries
management.

In this paper, we examine three strategies to managing a natural
population, such as a fishery, under such uncertainty. This problem has
a long history in both theory and practice, where seemingly innoccous
simplifying assumptions have led to a deep paradox in how managers
should respond to uncertainty. The precautionary principle (Kriebel et
al. 2001) would seem to suggest that the greater the uncertainty in our
models, the more cautious we should be in management policies. Yet a
well-known mathematical proof by Reed (1979) established that under
quite general conditions, the optimal strategy under uncertainty is no
different than under no uncertainty at all. Attempts to resolve this
have since only deepened this paradox, arguing that adding additional
sources of uncertainty, such as imperfect measurements, should lead to
optimal harvest that increases with greater uncertainty (e.g. Clark and
Kirkwood 1986; Sethi et al. 2005). Meanwhile, much of actual practice
has (perhaps fortunately, as we shall see) ignored these results in
favor of a more heuristic application of the precautionary principle.
Here, we show that this paradox can be resolved by uncovering previous
shortcuts and tackling the optimal management problem head-on, which we
can solve thanks to the computational efficiency of sophisticated
algorithms developed more recently in the field of robotics. This
approach allows us to reject this paradox largely in favor of the
precautionary principle. Further, our comparisons show that widely
recognized existing approaches in both theory and practice are not by
themselves sufficient to manage a population under substantial
uncertainty over the long term, but our recently borrowed approach from
robotics literature can under identical conditions sustain both higher
biomass and higher economic yields than those approaches. This
conclusion has important implications not only for fisheries but
eocsystem more broadly, underscoring the danger of not correctly
accounting for measurement uncertainty when making decisions, and
further to the significant opportunity to learn from fields such as
robotics when faced with complex and uncertain decisions.

@@

For simplicity we will assume measurement uncertainty is normally
distributed with standard deviation \(\sigma_m\),
\(\xi \sim \mathcal{N}(1, \sigma_m)\). Given an estimated population
model along with this (uncertain) measurement of the stock size, the
manager must choose the set of a harvest quotas \(\lbrace h_t \rbrace\)
for \(t \in [0, 1, \infty]\) that maximizes expected long term utility
of the stock:

\begin{equation}
\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot U(x_t, h_t) \right \}
\end{equation}

For convenience of presentation but without loss of generality, we will
assume this utility is simply equal to the harvest itself; in economic
terms, a fixed price for fish without cost on effort:

\begin{equation}
 U(x_t, h_t) = \min(h_t, x_t)
\end{equation}

where the \(\min\) function ensures no additional reward for attempting
to harvest more than the entire stock. Existing optimal control
techniques (Reed (1979)) prove these results are not dependent on the
particular details of this utility. Using this form allows a more direct
comparison to approaches such as maximum sustainable yield (MSY), that
do not include an economic model but merely maximize long-term catch. In
principle, utility could also reflect ecosystem services or other
intrinsic values for fish left in the sea and not merely the market for
consumption (Halpern et al. 2013).

The solution to this decision problem is conditional on a population
model for the underlying ecological process, describing probability the
system moves to state \(x_{t+1}\) given that it was in state \(x_t\) and
the manager selected a harvest \(h_t\), which we discuss below. Such a
model would typically be estimated and parameterized from available
data, often (but not necessarily) using a hierarchical Bayesian approach
that acknowledges the existence of uncertainty in both measurements and
process (e.g.~see Dichmont et al. 2016 for a discussion of how
uncertainties are frequently included in stock assessment models).
However, accounting for measurement uncertainty when going from data to
model does not mean we have also accounted for this uncertainty when
going from model to decision. For example, a harvest decision policy
based upon maximum sustainable yield simply depends on the stock size
which maximizes the expected growth rate of the model. Measurement
uncertainty in model estimation can indeed influence the position of
that maximum, but as we shall see, that strategy is only optimal in the
absence of any uncertainty.

There is one further subtlety arising in this seemingly simple problem
statement that we must address. Given our discrete-time formulation of
the recruitment process, it is necessary to decide if the measurement
\(y_t\) happens before or after harvest \(h_t\): that is do we: measure,
recruit, harvest, or measure, harvest, recruit? Following convention in
the optimal control literature (e.g. (Reed 1979; Clark 1990; Sethi et
al. 2005)), we will assume the latter; measurement occurs immediately
before harvest, followed by recruitment. This assumption has
occasionally been reversed to introduce some additional uncertainty from
stochastic recruitment immediately prior to setting the harvest quota
(e.g. Clark and Kirkwood 1986; Weitzman 2002), which appears to mimic
uncertainty in measurements but does not allow uncertainty to accumulate
since the measurement step still occurs without error.

The optimal strategy must think not only about the next year, but rather
considers the sequence of all future actions \({h_t}\), under all
possible realizations of the underlying model. This strategy must also
acknowledge that the manager will receive new information each year, and
may need to modify actions accordingly. It is this ability of the
decision problem to think ahead that allows it to tolerate short-term
costs (e.g.~reduced harvests in the current season) for larger future
payoffs. When Eq (1) is ignored, such optimal control problems are known
as Markov Decision Processes or MDPs. These are more commonly referred
to in the ecological literature by their solution method of Stochastic
Dynamic Programming (SDP; Marescot et al. 2013), and widely used in both
ecological management and animal behavior (Mangel 1985; Mangel and Clark
1988). The addition of measurement uncertainty in Eq (1) turns this
problem into a Partially Observed Markov Decision Process (POMDP), for
which solution methods have not been widely available for large problems
until more recently. We discuss these decision methods in more detail
below.

@@

It is important to remember that all three decision methods we compare
here will use this same model as given. In the case of numerical
simulations, we will simply pick reasonable fixed parameter values for
the model above. To illustrate the application to empirical data, we
will first estimate these parameters using a hierarchical Markov Chain
Monte Carlo (MCMC), and then use the resulting estimates to drive all
three decision models. As such, the details of how the model is
estimated are immaterial to the differences in performance between these
decision methods. In all cases, the model is considered fixed.
Technically, it would be possible to re-estimate the model parameters
given each new observation during the decision process. Adaptations of
the decision processes we consider here include this possibility for
learning about the model during the decisions. Some include the
possibility of taking knowingly bad actions, like suddenly increasing
harvest rates, to improve learning over parameters. We investigate these
in subsequent work (Memarzadeh and Boettiger 2018). For better
comparison with established theory, here, the model is considered fixed.
In the case of the simulations, the model is known without error (though
still stochastic in nature), such that additional observations cannot
improve the parameter estimates in any event. Despite that obvious
simplification, we shall illustrate how existing decision methods fail
to account for measurement uncertainty appropriately. Even with these
simple models, decisions are complex, and knowing the rules of the game
is not the same as knowing how to win.

@@

It is important to note that actual management strategies are quite
diverse. Our focus on comparisons TAC, MSY, and Constant Escapement (CE,
also refereed to as MEY, see below) reflects only the subset of current
practices. Another increasingly common approach to address uncertainty
in both process and models is known as Management Strategy Evaluation
(MSE, Smith 1994; Dorner et al. 2009; Bunnefeld et al. 2011; Punt et al.
2016). MSE performs forward simulations under a suite of pre-selected
candidate strategies to determine which strategy achieves the best
expected outcomes. Crucially, by limiting the analysis to a suite of
candidate strategies over a finite number of replicate simulations, this
approach is able to accommodate substantially greater complexity and
uncertainty than has been possible with more formal optimal control
approaches we discuss below. Optimal control solutions rely on dynamic
programming to consider all possible strategies (all possible sequences
of harvest under all possible realizations of stock dynamics). As such,
MSE can be viewed as an approximation to the optimal control solutions
considered below, a notion that has been formalized in the literature on
approximate dynamic programming (Nicol and Chadès 2011).

@@

In the face of a naturally fluctuating population, previous arguments
made by Gordon and Press (1954); Schaefer (1954) or Clark (1973) no
longer hold. As noted above, the optimal harvest of a population under a
model such as the stochastic Gordon Schaefer, Eq
\eqref{Gordon-Schaefer}, is an example of a Markov Decision Problem
(MDP) which must be solved with stochastic dynamic programming. Thanks
to a mathematical proof provided by Reed (1979), fisheries managers have
largely been able to bi-pass such computational effort. However, this
convenient result also opened the door to paradox of uncertainty that
would haunt ecological management for at least the next four decades.

@@

In practice, TAC-managed fisheries are more common that CE-managed
fisheries, at least in the United States. CE-based regulation is limited
primarily to salmon, species for which population estimates may be more
precise and thus closer to Reed's assumption. We are not aware of any
fishery that has put Clark's Paradox into practice, harvesting more in
response to increased uncertainty. As we shall illustrate, this is
fortunate; though approaches that properly account for both forms of
uncertainty can perform even better.

@@

Why does the introduction of measurement uncertainty make the decision
problem so much harder? Consider the thought experiment of managing a
fish stock where in each of the previous four years stock assessments
have put the population size at, say, 51, 54, 49, 51, (say, measured in
percent of carrying capacity, \(K\)). Under a constant escapement target
of 50, the harvest quota would be 1, 4, 0, 1, respectively. Note that
this control rule depends only on the measurement for the year in
question, which is assumed to be measured without error. If we then
measure a stock size of 75, an assumption that this measurement is made
without error would justify the sudden increase to a quota of 25. Yet if
we admit the possibility of measurement error, this sudden bump starts
to look very suspicious in light of all the previous observations.
Perfect measurements let us exploit the Markov property of the model --
the most recent measurement tells us all we need to know. In contrast,
the introduction of measurement uncertainty breaks this Markovian
assumption: a rational decision maker would suspect this 75 to be an
over-estimate based on prior observations. Under imperfect observations
introduced by Eq (1), we no longer have a Markov Decision Process (MDP)
over the observed state space. Rather, we have a Partially Observed
Markov Decision Process (POMDP). A POMDP policy cannot be expressed as a
policy function at all (much less as policy function with a single
constant-escapement target, as much previous work has assumed). Instead,
a POMDP policy can be thought of as depending not only on the most
recent observation, but the recent observation combined with a
\emph{prior belief} about what state the system is in, which itself is
determined by all past observations. In this manner, the POMDP can
consider how much weight to give the most recent observation based on
how it compares to prior beliefs.

@@

\hypertarget{simulations}{%
\subsection{Simulations}\label{simulations}}

We consider the average fish biomass across 100 replicate simulations of
the same stock dynamics (Eq 4) under three different management
strategies: constant escapement (CE), total allowable catch (TAC; equal
to 80\% MSY), and partially observed Markov decision process (POMDP)
management. Each suite of simulations is considered under three
sequentially higher measurement error regimes, from \(\sigma_m = 0\) (no
measurement errors), \(\sigma_m = 0.1\) (low error), and
\(\sigma_m = 0.15\) (moderate error), as indicated, while stochastic
recruitment (environmental noise) is set to a moderate
\(\sigma_g = 0.15\). Empirical estimates of measurement error in stock
sizes vary widely, with estimates often dependent upon methodological
assumptions. Early estimates have put possible error as high as 50\%
(Clark and Kirkwood 1986) and more recent analyses have suggested an
average of closer to 36\%, though with cases that may be much higher
(Ralston et al. 2011). Consequently, range between 0 to 15\% provides a
probably conservative estimate of typical unceratinty ranges, which we
show is nevertheless large enough to drive substantially negative
ecological and economic outcomes under management strategies that do not
appropriately account for that error over the long run.

@@

In the absence of either environmental noise or measurement error POMDP,
CE, and MSY would converge to stock at the \(B_{MSY}\), while TAC would
maintain the stock at a slightly higher level (see Supplemental
material, Figure S1). The introduction of stochastic growth has a
significant negative impact on the TAC strategy, with a mean biomass
significantly lower than \(B_{MSY}\) (first panel, Figure \ref{sims};
this impact is even more severe for MSY, see Figure S2). Without
measurement error, the CE and POMDP strategies are nearly identical with
both approximately maintaining the stock at \(B_{MSY}\) despite the
significant environmental stochasticity. As measurement error increases
in the subsequent panels, CE and TAC strategies perform increasingly
poorly, while the POMDP continues to maintain the average stock close to
\(B_{MSY}\). Notably, CE is significantly more impacted by measurement
uncertainty than TAC, with CE averaging even lower biomass than TAC
under moderate measurement uncertainty. These results confirm that the
precautionary approach represented by the TAC does indeed prove more
robust to the problem of measurement error than the optimization
solution represented by CE, but is less robust to environmental error.
In contrast, the POMDP approach successfully handles both sources of
uncertainty, maintaining higher stocks, near \(B_{MSY}\) level, and
producing the highest yield.

@@

Using modern algorithms we have been able to both crack the nut of
measurement uncertainty in the optimal management of ecological
populations such as marine fisheries, debunking a long-standing paradox
created by previous literature which has argued that increased
measurement uncertainty should be met with equal or larger harvests
(e.g. Reed 1979; Clark and Kirkwood 1986; Sethi et al. 2005). By using
powerful algorithms developed first the robotics literature to solve
POMDPs (Kurniawati et al. 2008), we have been able to tackle this
optimal control problem directly rather than resorting to the
simplifying assumptions used in previous work. Our solution confirms the
intuition of the precautionary principle in a quantitatively precise
form: under most (but not all) conditions, \emph{greater uncertainty
results in smaller optimal harvests}. Despite frequently proposing
smaller harvests than existing approaches, the POMDP solution achieves
higher long-term economic value as well as higher average stock sizes
over the long term.

This lesson has clear relevance in the context of fisheries management.
Here, practice is already wiser than theory, where the optimal control
based on the CE results of Clark (1973) and Reed (1979) is limited to
fisheries such as salmon where estimates in stock size may be less
significant than in purely marine species. To our knowledge, no fishery
has intentionally put the more aggressive harvest solutions of Clark and
Kirkwood (1986) and Sethi et al. (2005) into practice. Yet this does not
lift the specter of these methods, where CE-based predictions feature
prominently in global fisheries analyses as the economically optimal
strategy and the fastest route to recovery (Costello et al. 2016;
Burgess et al. 2018). Our results raise serious questions about the
optimistic forecasts predicted under these CE-based policies. Meanwhile,
MSY remains the dominate standard international law (Mace 2001), and
many well-managed US fisheries rely on a version of the TAC based
precautionary adjustment to MSY to comply with catch limits under
Magnuson-Stevens Act. Our results show that these strategies are less
sensitive to measurement error, but not immune. Comparisons against more
heuristic decision methods are more difficult: management strategy
evaluation (MSE; Smith 1994; Punt et al. 2016) will depend on precisely
what management scenarios are considered. Under sufficient uncertainty,
both the existing optimal and precautionary approaches can lead to
declines in biomass and economic returns which could be avoided by
POMDP-based management.

This conclusion has important broader implications for ecological
management more generally. Decision making under uncertainty is a
central challenge for ecology and conservation biology, which frequently
pits optimal control based approaches against more heuristic approaches
such as resilience thinking, scenario planning, and the precautionary
principle (Polasky et al. 2011). With economists frequently favoring the
precise, stylized approach of optimal control and ecologists embracing
the realities of greater complexity and uncertainty echoed in more
heuristic approaches, this division can sometimes appear to pit economic
outcomes against ecological objectives. That perspective is echoed in
the paradox discussed here, in which economic optimization ignoring
uncertainty has argued in cold mathematical logic for a rejection of the
precautionary principle. Our results demonstrate that this divide is
artificial. With sufficiently powerful algorithms we can go some ways to
realizing the call of Fischer et al. (2009) to combine more ecologically
realistic assumptions such as measurement error with optimal control
approaches in ecological management. Both sides are vindicated in our
solution: economic optimization under POMDP does not contradict the
precautionary principle, while under POMDP, optimal control can also
achieve better ecological outcomes than blunt precautionary rule
(i.e.~TAC, which reduced all harvests by 20\% below MSY target). As
Fischer et al. (2009) predicted, unifying optimal control with more
resilience-inspired approaches is win-win.

If accounting for measurement uncertainty can resolve previous paradoxes
and divides, it also raises additional challenges. The approach
discussed here is vulnerable both to the charge of being too simple to
be realistic while also being too complicated to be feasible. These
issues are as easily recognized in other ecological management contexts,
from harvests of wild game or forestry to pest outbreaks, (e.g. Richards
et al. 1999; Shea 1998; Blackwood et al. 2010; Shea et al. 2014; Dee et
al. 2017; Iacona et al. 2017; Li et al. 2017) as in fisheries, where
ecological complexity is invariably high and management capacity
invariably limited. We address each of these in turn.

Are the dynamics considered here too simple? Our underlying model is
undoubtedly much simpler than many of the age or stage structured models
used in modern fisheries management -- aspects whose importance to
ecological dynamics have been established well beyond fisheries as well.
Yet it is critical to bear in mind that this stylized one-dimensional
model is used only to compare the relative performance of the different
decision strategies considered here (TAC, CE, and POMDP) in order to
illustrate the potential importance of uncertainty in measurement. There
is little reason to believe CE or TAC would perform better under more
complex cases such as age-structured models that only further deviate
from the assumptions under which they were derived (e.g. Holden and
Conrad 2015). POMDP approaches are in no way incompatible with more
complex models, whether that be age structure or even more ambitious
non-parametric representations defined directly from data using Gaussian
Processes or other machine learning techniques (e.g. Boettiger et al.
2015). Simple models have always served as tools for comparison and
intuition, and serve the same illustrative role here.
